---
title: "sqlite in R"
author: "Dipak Bange"
date: "2/6/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Uncomment the code below once to install necessary libraries, then comment again. You only need it once.

```{r}
#install.packages("RSQLite")
#install.packages("DBI")
```

Load the RSQLite Library

```{r}
library(RSQLite)
library(DBI)
```

## Creating Database

### Step 1.

Creating connection and database. We call database CarsDB.db and we are going to use mtcars dataset

```{r}
conn <- dbConnect(RSQLite::SQLite(), dbname = "CarsDB.db")
```

Note - when you save your file, your should see CarsDB.db in your directory.

### Step 2.

We are cleaning data mtcars

```{r}
data("mtcars")
head(mtcars) # this is what it looks like with car names as row names
# row names are like row indices
mtcars$car_names <- rownames(mtcars) # creating a new column from row names
rownames(mtcars) <- c() # table row names are converted back to numeric indices (no names)
head(mtcars)
```

### Step 3.

Insert data to database. Note - once you done with Step 3, you need to add eval=FALSE inside the curly brackets `{r eval=FALSE}`

```{r eval = FALSE}
# Write the mtcars dataset into a table names mtcars_data
 dbWriteTable(conn, "cars_data", mtcars)
# List all the tables available in the database
dbListTables(conn) 
```

### Step 4.

Append tables. Note - once you done with Step 4, you need to add eval=FALSE inside the curly brackets `{r eval=FALSE}`

```{r}
# Create toy data frames
car <- c('Camaro', 'California', 'Mustang', 'Explorer')
make <- c('Chevrolet','Ferrari','Ford','Ford')
df1 <- data.frame(car,make)
car <- c('Corolla', 'Lancer', 'Sportage', 'XE')
make <- c('Toyota','Mitsubishi','Kia','Jaguar')
df2 <- data.frame(car,make)
# Add them to a list
dfList <- list(df1,df2)
# Write a table by appending the data frames inside the list
for(k in 1:length(dfList)){
    dbWriteTable(conn,"Cars_and_Makes", dfList[[k]], append = TRUE)
}
# List all the Tables
dbListTables(conn)
```

### Step 5. Queries

Query 1. Gather data from Cars_and_Makes

```{r}
dbGetQuery(conn, "SELECT * FROM Cars_and_Makes")
```

Query 2: Gather the first 10 rows in the cars_data table

```{r}
dbGetQuery(conn, "SELECT * FROM cars_data LIMIT 10")
```

Query 3. Get the car names and horsepower of the cars with 8 cylinders

```{r}
dbGetQuery(conn,"SELECT car_names, hp, cyl FROM cars_data WHERE cyl = 8;")
```

Query 4.Get the car names and horsepower starting with M that have 6 or 8 cylinders

```{r}
dbGetQuery(conn,"SELECT car_names, hp, cyl FROM cars_data
                 WHERE car_names LIKE 'M%' AND cyl IN (6,8)")

```

Query 5. Get the average horsepower and mpg by number of cylinder groups

```{r}
dbGetQuery(conn,"SELECT cyl, AVG(hp) AS 'average_hp', AVG(mpg) AS 'average_mpg' FROM cars_data
                 GROUP BY cyl
                 ORDER BY average_hp")
```

### Step 4. Storing

```{r}
avg_HpCyl <- dbGetQuery(conn,"SELECT cyl, AVG(hp) AS 'average_hp'FROM cars_data
                 GROUP BY cyl
                 ORDER BY average_hp")
avg_HpCyl
class(avg_HpCyl)
```

### Step 5. Parametrized Queries

```{r}
mpg <-  18
cyl <- 6
Result <- dbGetQuery(conn, 'SELECT car_names, mpg, cyl FROM cars_data WHERE mpg >= ? AND cyl >= ?', params = c(mpg,cyl))
Result
```

### Step 6. Inserting and Deleting

```{r}
# Insert the data for the Mazda RX4. This will also ouput a 1
dbExecute(conn, "INSERT INTO cars_data VALUES (21.0,6,160.0,110,3.90,2.620,16.46,0,1,4,4,'Mazda RX4')")

# Delete the column belonging to the Mazda RX4. You will see a 1 as the output.
dbExecute(conn, "DELETE FROM cars_data WHERE car_names = 'Mazda RX4'")

dbGetQuery(conn, "SELECT * FROM cars_data")
```

### Step 7. Disconnecting

Once you are done operating with your SQLite database within R, it is important to call the function `dbDisconnect()`. This ensures that we release any resources that the database connection has been using, which is always a good practice.

```{r}
dbDisconnect(conn)
```
